---
title: "Stationary process or not"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tibble)
library(ggplot2)
library(purrr)
library(reshape2)

knitr::opts_chunk$set(echo = FALSE)
```

## Welcome

In the exercises you will see realizations of time series. In each exercise, based on the figure, you need to decide if the time series is stationary or not. Remember that a time series $\{X_t : t \in \mathbb{Z}\}$ is stationary if

1. $\mathbb E |X_t|^2 < \infty$ for all $t\in\mathbb{Z}$,

2. $\mathbb E X_t = m$ for all $t\in\mathbb{Z}$ and

3. $\mathrm{Cov}(X_r, X_s) = \mathrm{Cov}(X_{r + t}, X_{s + t})$ for all $r,s,t\in\mathbb{Z}$.

Covariance between two random variables $X$ and $Y$ is defined as $\mathrm{Cov}(X, Y) = \mathbb E((X - \mathbb E X)(Y - \mathbb E Y))$. Covariance tells if there is linear dependence between two random variables:

- $\mathrm{Cov}(X, Y) = 0 \Rightarrow$ $X$ and $Y$ are linearly independent.

- $\mathrm{Cov}(X, Y) > 0 \Rightarrow$ $X$ and $Y$ are positively correlated.

- $\mathrm{Cov}(X, Y) < 0 \Rightarrow$ $X$ and $Y$ are negatively correlated.

## Exercise 1: Carbon Dioxide Measurements

Below time series represents montly carbon dioxide measurements of the Mauna Loa volcano in Hawaii from 1959 to 1997 (one measurement per month, 468 observations).

```{r carbon-figure}
co2 <- tibble(value = as.vector(co2), time = zoo::as.Date(time(co2)))
ggplot(co2, aes(x = time, y = value)) +
  geom_line() +
  scale_x_date(date_labels = "%m/%Y", date_breaks = "5 year") +
  labs(x = "Time", y = "Value") +
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r carbon-question}
question("Is the time series stationary?",
         answer("no", correct = TRUE),
         answer("yes"),
         message = "The time series is not stationary. There is an upward trend,
         which violates Assumption 2. Also, there seems to be seasonality, which
         violates Assumption 2 as well.")
```

## Exercise 2: Simulated Data from $AR(1)$ Process

Let $\varepsilon_t \sim N(0,1)$, $t \in\{1, 2, \ldots, 1000\}$, be independent and identically distributed random variables from the standard normal distribution. Below time series is simulated from the so called $\mathrm{AR}(1)$ process (AR = autoregressive),
\begin{equation*}
X_t =
  \begin{cases}
    0, & t = 0 \\
    \theta X_{t-1} + \varepsilon_t, & t\in\{1, 2, \ldots, 1000\},
  \end{cases}
\end{equation*}
when we set $\theta = 0.5$. 

```{r ar-figure}
n <- 1000
theta <- 0.5
epsilon <- rnorm(n)
x <- rep(0, n + 1)
for (i in 2:length(x)) {
  x[i] <- theta * x[i - 1] + epsilon[i - 1]
}

tibble(value = x, time = 0:n) |>
  ggplot(aes(x = time, y = value)) +
  geom_line() +
  labs(x = "Time", y = "Value") +
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_blank())
```

```{r ar-question}
question("Is the time series stationary?",
         answer("no"),
         answer("yes", correct = TRUE),
         message = "The time series is stationary. There is no trend or
         seasonality, which suggests that the stationarity assumption might
         hold. It is quite hard to see in this figure alone, but also the
         Assumption 3 is satisfied. In order to check Assumption 3 we need
         more tools such as the autocorrelation function.")
```

## Exercise 3: Simple Random Walk

Let $\varepsilon_t$, $t \in\{1, 2, \ldots, 100\}$, be independent random variables such that $\mathbb{P}(\varepsilon_t = -1) = \mathbb{P}(\varepsilon_t = 1) = 1/2$. Below figure includes 10 realizations of the simple random walk,
\begin{equation*}
X_t =
  \begin{cases}
    0, & t = 0 \\
    X_{t-1} + \varepsilon_t, & t\in\{1, 2, \ldots, 100\},
  \end{cases}
\end{equation*}

```{r simple-figure}
m <- 10
n <- 100
seed <- 123

set.seed(seed)
paths <- rerun(m, sample(c(-1, 1), n, replace = TRUE)) |>
  map(~ c(0, .x)) |>
  map(cumsum) |>
  set_names(letters[1:m]) |>
  as_tibble() |>
  dplyr::mutate(index = 0:n)

colors <- c(rep("gray80", m - 1), "black")

melt(paths, id.vars = "index", variable.name = "path", value.name = "value") |>
  ggplot(aes(x = index, y = value, group = path, color = path)) +
  geom_point() +
  geom_line() +
  labs(x = "Time", y = "Value") +
  scale_color_manual(values = colors) +
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.background = element_blank(),
        legend.position = "none")
```

```{r simple-question}
question("Is the time series stationary?",
         answer("no", correct = TRUE),
         answer("yes"),
         message = "The time series is not stationary. Assumption 3 is violated
         since variance blows up as time increases. This phenomenon is visible
         on the above figure.")
```
